---
title: "April 2022 Waste Audit"
author: ''
subtitle: Duke Sustainability Task Force
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

to do: 
- fix can_number -> Location tag for 28-30 (why showing up NA?)
- t-distributions, change the degree of freedom
- slices of the data taken for ANY graph or table
- compare feb to apr
- compare 2022 to 2021

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(infer)
library(car)
library(stringr)
library(robotstxt)
library(rvest)
library(httr)
library(tinytex)
library(sf)
```

```{r upload-and-clean-data}
waste_audit <- april2022auditdata %>% 
  mutate(total_piece_volume = pieces_trash + pieces_recycle + pieces_compost, 
        total_can_weight = weight_trash_lbs + weight_recycle_lbs + weight_compost_lbs, 
        prop_pieces_trash = pieces_trash / total_piece_volume,
        prop_pieces_recycle = pieces_recycle / total_piece_volume,
        prop_pieces_compost = pieces_compost / total_piece_volume,
        prop_trash_weight = weight_trash_lbs / total_can_weight,
        prop_recycle_weight = weight_recycle_lbs / total_can_weight,
        prop_compost_weight = weight_compost_lbs / total_can_weight)
```

```{r average-piece-weight}
waste_audit <- waste_audit %>% 
  mutate(restaurant = case_when(
    can_number %in% c(23, 24, 25, 26, 28, 27) ~ "Zwelis",
    can_number %in% c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22) ~ "Loop",
    can_number %in% c(38, 39, 40, 39, 31, 47, 48, 45, 41, 35, 36, 37, 34, 32, 33, 49, 30, 29) ~ "Plaza"
  ))
```

#Hailey, cans 29, 30, and 28 are showing up NA under can_number... FIX

# Methodology

## Content Comparisons
- average weight of a piece of trash, composting, and recycling item
```{r average-piece-weight}
waste_audit %>% 
  summarize(total_trash_pieces = sum(pieces_trash),
            total_recycle_pieces = sum(pieces_recycle),
            total_compost_pieces = sum(pieces_compost),
            total_trash_weight = sum(weight_trash_lbs), 
            total_recycle_weight = sum(weight_recycle_lbs),
            total_compost_weight = sum(weight_compost_lbs),
            average_trash_weight = total_trash_weight / total_trash_pieces,
            average_recycle_weight = total_recycle_weight / total_recycle_pieces,
            average_compost_weight = total_compost_weight / total_compost_pieces)
```

- comparative starting volume of pieces/weight between trash, recycle, and
compost streams
```{r starting-volumes}
waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(avg_starting_weight = mean(total_can_weight),
            avg_starting_pieces = mean(total_piece_volume))
```

```{r bar-plot-original-stream-piece-comparison}
data <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(total_trash = sum(pieces_trash),
            total_recycle = sum(pieces_recycle),
            total_compost = sum(pieces_compost)) %>% 
  select(total_trash, total_recycle, total_compost)
data <- as.matrix(data)
colnames(data) <- c("trash", "recycle", "compost")
rownames(data) <- c("Compost", "Recycle", "Trash")
color <- c("black", "#3265ab", "#2ca816")
barplot(t(data), beside = FALSE, col = color, ylim = c(0, 2500),
        main = "Actual Waste Composition of Trash, Recycling, and Compost (in pieces)", xlab = "Original Stream Designation", ylab = "Total Pieces")
  legend("topleft", colnames(data), pch = 15, col = color, horiz = T, bty = "n")
# source: https://stackoverflow.com/questions/61506408/grouped-barchart-in-r-with-4-variables
```

## Quantifying Missorted Waste
```{r missorted}
waste_audit <- waste_audit %>% 
  mutate(perc_missorted_c = case_when(original_stream == "trash" ~ (pieces_recycle + pieces_compost) / total_piece_volume, original_stream == "recycling" ~ (pieces_trash + pieces_compost) / total_piece_volume, original_stream == "compost" ~ (pieces_trash + pieces_recycle) / total_piece_volume),
         missorted_count = case_when(original_stream == "trash" ~ (pieces_recycle + pieces_compost), original_stream == "recycling" ~ (pieces_trash + pieces_compost), original_stream == "compost" ~ (pieces_trash + pieces_recycle)),
         sorted_count = total_piece_volume - missorted_count)
```

```{r bar-plot-original-stream-percent}
makeup <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(average_trash = mean(prop_pieces_trash),
            average_recycle = mean(prop_pieces_recycle),
            average_compost = mean(prop_pieces_compost)) %>% 
  select(average_trash, average_recycle, average_compost)
makeup <- as.matrix(makeup)
colnames(makeup) <- c("trash", "recycle", "compost")
rownames(makeup) <- c("Compost", "Recycle", "Trash")
color <- c("black", "#3265ab", "#2ca816")
barplot(t(makeup), beside = FALSE, col = color, ylim = c(0, 1.1),
        main = "Percent Actual Waste Composition in Trash, Recycling, and Compost", xlab = "Original Stream Designation", ylab = "Proportion")
legend("topleft", colnames(makeup), pch = 15, col = color, horiz = T, bty = "n")
```

- percent of total waste that is missorted

```{r missorted-total-waste}
pie_total <- waste_audit %>% 
  summarize(sorted = sum(sorted_count),
            missorted = sum(missorted_count))
pie_total
slices <- c(1301, 1636)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Overall Contamination for All Streams")
# source: https://www.statmethods.net/graphs/pie.html
```

- Bootstrap Sample OVERALL WASTE

bootstrap sample

```{r bootstrap-sample-with-compost}
set.seed(201)
boot_missorted <- waste_audit %>% 
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-with-compost}
ci_boot_missorted <- get_ci(boot_missorted, level = 0.95)
ci_boot_missorted
```

bootstrap distribution

```{r bootstrap-vis-with-compost}
visualize(boot_missorted) +
  shade_ci(ci_boot_missorted) +
  labs(title = "Simulated Bootstrap Distribution of Overall Missorted Waste", x = "Sample Mean Missorted Overall Waste", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-with-compost, eval = FALSE}
mean(waste_audit$perc_missorted_c) + c(-1, 1) * 2.05 * sd(waste_audit$perc_missorted_c) / sqrt(44)
```

tdist: 0.4984849, 0.5994991

#change square root above because i think that it is wrong

-percent of trash and recycling that is missorted (so we can compare last year)

```{r missorted-trash-plus-recycling}
pie_tandr <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(sorted_count),
            missorted = sum(missorted_count)) %>%
  slice(2:3)
pie_tandr <- pie_tandr %>%
  summarize(sorted = sum(sorted),
            missorted = sum(missorted))
pie_tandr
slices <- c(1055, 1424)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Overall Contamination in both Trash and Recycling")
```

- Bootstrap Sample TRASH + RECYCLING ONLY

bootstrap sample

```{r bootstrap-sample-trash-plus-recycling}
boot_table_tr <- waste_audit %>%
  slice(1:8, 11:19, 22:24, 27:30, 32, 34:36, 38:41)
set.seed(201)
boot_missorted_tr <- boot_table_tr %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-trash-plus-recycling}
ci_boot_missorted_tr <- get_ci(boot_missorted_tr, level = 0.95)
ci_boot_missorted_tr
```

bootstrap distribution

```{r bootstrap-vis-trash-plus-recycling}
visualize(boot_missorted_tr) +
  shade_ci(ci_boot_missorted_tr) +
  labs(title = "Simulated Bootstrap Distribution of Overall Trash & Recycling", x = "Sample Mean Missorted Trash & Recycling", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-trash-plus-recycling, eval = FALSE}
mean(boot_table_tr$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_tr$perc_missorted_c) / sqrt(32)
```

tdist: 0.5192801, 0.6199768

#change square root above because i think that it is wrong

- percent of trash that is missorted

```{r missorted-trash}
pie_trash <- waste_audit %>% 
  mutate(missorted = pieces_recycle + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_trash),
            missorted = sum(missorted)) %>%
  slice(3)
pie_trash
slices <- c(697, 1137)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Trash")
```

- Bootstrap Sample TRASH ONLY

bootstrap sample

```{r bootstrap-sample-trash}
boot_table_trash <- waste_audit %>%
  slice(1:6, 11:17, 22:24, 27:29, 32, 35, 38:41)
set.seed(201)
boot_missorted_trash <- boot_table_trash %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-trash}
ci_boot_missorted_trash <- get_ci(boot_missorted_trash, level = 0.95)
ci_boot_missorted_trash
```

bootstrap distribution

```{r bootstrap-vis-trash}
visualize(boot_missorted_trash) +
  shade_ci(ci_boot_missorted_trash) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Trash", x = "Sample Mean Missorted Trash", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-trash, eval = FALSE}
mean(boot_table_trash$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_trash$perc_missorted_c) / sqrt(25)
```

tdist: 0.5246678, 0.6368954

#change square root above because i think that it is wrong

- percent of recycling that is missorted
    
```{r missorted-recycle}
pie_recycle <- waste_audit %>% 
  mutate(missorted = pieces_trash + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_recycle),
            missorted = sum(missorted)) %>%
  slice(2)
pie_recycle
slices <- c(358, 287)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Recycle")
```

- Bootstrap Sample RECYCLING ONLY

bootstrap sample

```{r bootstrap-sample-recycling}
boot_table_recycling <- waste_audit %>%
  slice(7:8, 18:19, 30, 34, 36)
set.seed(201)
boot_missorted_recycling <- boot_table_recycling %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-recycling}
ci_boot_missorted_recycling <- get_ci(boot_missorted_recycling, level = 0.95)
ci_boot_missorted_recycling
```

bootstrap distribution

```{r bootstrap-vis-recycling}
visualize(boot_missorted_recycling) +
  shade_ci(ci_boot_missorted_recycling) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Recycling", x = "Sample Mean Missorted Recycling", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-recycling, eval = FALSE}
mean(boot_table_recycling$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_recycling$perc_missorted_c) / sqrt(7)
```

tdist: 0.4056565, 0.6443756

#change square root above because i think that it is wrong

- percent of compost that is missorted
    
```{r missorted-compost}
pie_compost <- waste_audit %>% 
  mutate(missorted = pieces_trash + pieces_recycle) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_compost),
            missorted = sum(missorted)) %>% 
  slice(1)
pie_compost
slices <- c(246, 212)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Compost")
```

- Bootstrap Sample COMPOST ONLY

bootstrap sample

```{r bootstrap-sample-compost}
boot_table_compost <- waste_audit %>%
  slice(9, 10, 20, 21, 25, 26, 31, 33, 37, 45, 47, 48)
set.seed(201)
boot_missorted_compost <- boot_table_compost %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-compost}
ci_boot_missorted_compost <- get_ci(boot_missorted_compost, level = 0.95)
ci_boot_missorted_compost
```

bootstrap distribution

```{r bootstrap-vis-compost}
visualize(boot_missorted_compost) +
  shade_ci(ci_boot_missorted_compost) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Compost", x = "Sample Mean Missorted Compost", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-compost, eval = FALSE}
mean(boot_table_compost$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_compost$perc_missorted_c) / sqrt(12)
```

tdist: 0.3526503, 0.6163557

#change square root above because i think that it is wrong

## Campus Compost
- percent of all waste that could be compost
    
```{r compost-in-total-waste}
compostable_count <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable))
compostable_count
compostables <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable))
compostables
slices <- c(1200, 1737)
lbls <- c("compostable", "noncompostable")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Total Waste that is Actually Compostable")
```

 - percent of trash that could be compost

```{r compost-in-trash}
compostable_trash <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable)) %>% 
  slice(3)
compostable_trash
slices <- c(837, 997)
lbls <- c("compostable", "noncompostable")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Trash that is Actually Compostable")
```

 - percent of recycling that could be compost

```{r compost-in-recycling}
compostable_recycle <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable)) %>% 
  slice(2)
compostable_recycle
slices <- c(117, 528)
lbls <- c("compostable", "noncompostable")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Recycling that is Actually Compostable")
```

 - percent of compost that is supposed to be composted

```{r compost-in-compost}
compostable_compost <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable)) %>% 
  slice(1)
compostable_compost
slices <- c(246, 212)
lbls <- c("compostable", "noncompostable")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Compost that is Actually Compostable")
```

- volume of compostables
    
```{r compost-volume-values}
volume_bar <- waste_audit %>% 
  summarize(total_pieces_compost = sum(pieces_compost),
            total_weight_compost = sum(weight_compost_lbs),
            total_pieces_audited = sum(total_piece_volume),
            total_weight_audited = sum(total_can_weight),
            perc_pieces_compost = total_pieces_compost / total_pieces_audited * 100,
            perc_weight_compost = total_weight_compost / total_weight_audited * 100)
makeup <- as.matrix(volume_bar)
volume_bar <- volume_bar %>% 
  select(perc_pieces_compost, perc_weight_compost)
colnames(volume_bar) <- c("Pieces", "Weight (lbs)")
color <- c("#2ca816", "#8cd9b3")
barplot(t(volume_bar), beside = TRUE, col = color, ylim = c(0, 100),
        main = "Percent of Total Waste that is Actually Compostable by Piece Count and Weight", xlab = "Method of Quantifying Waste", ylab = "Percent")
legend("topleft", colnames(volume_bar), pch = 15, col = color, horiz = T, bty = "n")
```

## Differences in Location... INCOMPLETE SECTION BECAUSE OF ERROR IN LOCATION CODE
```{r location-contamination-table}
location_contamination <- waste_audit %>%
  group_by(restaurant) %>%
  summarize(sorted = sum(sorted_count),
    missorted = sum(missorted_count))
location_contamination
```

contamination at Zweli's Café

```{r zwelis-contamination}
slices <- c(172, 254)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Contamination at Zweli's Café")
```

contamination at BC

```{r plaza-contamination}
slices <- c(589, 938)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Contamination on BC Plaza")
```

contamination at the loop

```{r loop-contamination}
slices <- c(409, 668)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Contamination at The Loop")
```

## Comparison Between February and April

## Comparison Between 2021 and Feb 2022 Results

# DID NOT LOOK AT TABLE BELOW AND ALSO DO MORE

```{r 2021-feb2022}
con <- data.frame(year = c("2021", "2022"),
  contamination_rate = c(74, 61))
ggplot(con) +
    geom_bar(aes(x=year, y=contamination_rate), stat="identity", fill="#2ca816", width = 0.5) +
  xlab("Year") +
  ylab("Contamination Rates") +
    ggtitle("Yearly Contamination Rate Comparison") +
    ylim(0, 100) +
  theme_minimal()
```

# Discussion

-	error: human error in sorting
-	error: some contamination spreads (e.g. containers filled with liquid)
-	error: varying time between locations on how long the waste receptacles were allowed to fill up between last emptying time and time of collection
-	discuss time of day and year that the data was collected
-	note: this does not account for all waste; waste disposed of in dorms, WU conveyor belt, BC, library, quad trash bins, gym (Red Mango), East Campus (Marketplace, dorms)
-	random sampling from the Plaza Big Bellies --- we got through all of the trash bags that we could with no particular order in mind
-	common contaminants: clear plastic lids (some are compostable and others are recyclable)
-	outstanding contaminants: 2 reusable bowls from WU
