---
title: "Compare February to April 2022 Waste Audits"
author: "Duke Sustainability Task Force"
date: "2022-09-02"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(infer)
library(car)
library(stringr)
library(robotstxt)
library(rvest)
library(httr)
library(tinytex)
library(readr)
```

```{r read-csv}
compareauditdata <- read_csv('Data Feb-Apr 2022 - Sheet1.csv')
View(compareauditdata)
```

```{r clean-data}
waste_audits <- compareauditdata %>% 
  mutate(total_piece_volume = pieces_trash + pieces_recycle + pieces_compost, 
        total_can_weight = weight_trash_lbs + weight_recycle_lbs + weight_compost_lbs, 
        prop_pieces_trash = pieces_trash / total_piece_volume,
        prop_pieces_recycle = pieces_recycle / total_piece_volume,
        prop_pieces_compost = pieces_compost / total_piece_volume,
        prop_trash_weight = weight_trash_lbs / total_can_weight,
        prop_recycle_weight = weight_recycle_lbs / total_can_weight,
        prop_compost_weight = weight_compost_lbs / total_can_weight)
```

```{r average-piece-weight}
waste_audits %>% 
  summarize(total_trash_pieces = sum(pieces_trash),
            total_recycle_pieces = sum(pieces_recycle),
            total_compost_pieces = sum(pieces_compost),
            total_trash_weight = sum(weight_trash_lbs), 
            total_recycle_weight = sum(weight_recycle_lbs),
            total_compost_weight = sum(weight_compost_lbs),
            average_trash_weight = total_trash_weight / total_trash_pieces,
            average_recycle_weight = total_recycle_weight / total_recycle_pieces,
            average_compost_weight = total_compost_weight / total_compost_pieces)
```

```{r missorted}
waste_audits <- waste_audits %>% 
  mutate(perc_missorted_c = case_when(original_stream == "trash" ~ (pieces_recycle + pieces_compost) / total_piece_volume, original_stream == "recycling" ~ (pieces_trash + pieces_compost) / total_piece_volume, original_stream == "compost" ~ (pieces_trash + pieces_recycle) / total_piece_volume),
         missorted_count = case_when(original_stream == "trash" ~ (pieces_recycle + pieces_compost), original_stream == "recycling" ~ (pieces_trash + pieces_compost), original_stream == "compost" ~ (pieces_trash + pieces_recycle)),
         sorted_count = total_piece_volume - missorted_count)
```

```{r calculate-totals}
waste_audits %>% 
  group_by(month) %>% 
  summarize(sumtrash = sum(pieces_trash),
            sumrecycle = sum(pieces_recycle),
            sumcompost = sum(pieces_compost))

1118 + 265 + 1647

1018 + 719 + 1200
```

3030 pieces overall in February
2937 pieces overall in April

```{r}
waste_audits %>% 
  filter(original_stream == "trash") %>% 
  group_by(month) %>% 
  summarize(sum = sum(pieces_trash))
```

```{r}
waste_audits %>% 
  filter(original_stream == "recycling") %>% 
  group_by(month) %>% 
  summarize(sum = sum(pieces_recycle))
```

```{r missorted}
waste_audits %>% 
  filter(original_stream == "compost") %>% 
  group_by(month) %>% 
  summarize(sum = sum(pieces_compost))
```

```{r}
783 + 105 + 282
697	+ 358 + 246
```

1170 correctly sorted pieces in February
1301 correctly sorted pieces in April

```{r overall-contamination}
prop.test(x = c(1170, 1301), n = c(3030, 2937))
```

Overall contamination
38.61% -> 44.30%  with a p-value < 0.05

break down by category -- trash, recycling, and compost
recycling definitely driving results
could do this for recycling specifically
finding what percent of each category were correct per month (filter by type of item)

```{r trash-contamination}
prop.test(x = c(783, 697), n = c(1118, 1018))
```

Trash contamination
Not a significant change month to month for trash because p-value = 0.46 > 0.05.

```{r recycle-contamination}
prop.test(x = c(105, 358), n = c(265, 719))
```

Recycling contamination
39.62% -> 49.79%  with a p-value 0.0057 < 0.05

```{r compost-contamination}
prop.test(x = c(282, 246), n = c(1647, 1200))
```

Compost contamination
17.12% -> 20.50%  with a p-value 0.025 < 0.05

1 graphic with side by side
2 bars trash (Feb & Apr), 2 bars recycling, 2 bars compost, 2 overall

Results suggestion campaign was successful -- focus monre attention on inconclusive/negative results.























```{r stat}
#stat <- waste_audits %>%
#specify(perc_missorted_c ~ month) %>%
#calculate(stat = "diff in means", order = c("February", "April"))
```

```{r null-dist}
#null_dist <- waste_audits %>%
#specify(perc_missorted_c ~ month) %>%
#hypothesize(null = "independence") %>%
#generate(reps = 10000, type = "permute") %>%
#calculate(stat = "diff in means", order = c("February", "April"))
```

```{r visualize}
#visualize(null_dist) +
  #shade_p_value(obs_stat = stat, direction = "two-sided")
```

```{r p-value}
#null_dist %>%
  #get_p_value(obs_stat = stat, direction = "two-sided") 
```
